An Introduction to Return Oriented Programming

### 2023-02-17 paper.tex
changes

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
BIBLIOGRAFIE NICHT VERGESSEN
Return Oriented Programming is a type of buffer overflow attack that has been published in 2007 and ever since has become a widely known buffer overflow technique. It has been developed to circumvent the NX-BIT protection that protects the stack from being executed. At the time of writing this paper modern techniques like Stack Carnaries and ASLR prevent these attacks from being practical but there are millions of running systems using old hard-, firm- and software that is possibly vulnerable to these kinds of buffer overflow attacks. Return Oriented Programming is based on chaining return addresses to code just before a return and therefor allowing almost arbitrary code segments to be chained.

\section{Gadgets}
\label{sec:main}
\paragraph{Introduction}
On the x86 architecture the \Verb+ret+ instruction is defined to pop the return instruction pointer from the stack into the \Verb+eip+ register and redirect code execution to that memory address. By chaining addresses of instructions that end on a return and injecting them 
Gadgets are code segments that sit before a \Verb+ret+ instruction, these assembly instructions can be chained arbitrarily

@@

A gadget can be found by searching for \Verb+0xC3+ Bytes in the program. The instructions before then represent the code we can use, for that we need the address of the gadget. It is possible this manually using tools like \Verb+objdump+, \Verb+hexdump+ or use one of the many tools available, to name a few there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+. For this paper i will be using \Verb+ROPgadget+ since i found it easy to use and fast. \Verb+ROPgadget+ can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}. The gadgets can be extracted from the file using the following command~\cref{dumpallgadgets}. We can then use regular expressions to search for the gadgets that we need.

@@

These are only 10 Lines out of the 8244 lines found by the tool though and i purposefully filtered out some good and bad ones for demonstration. It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB. Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state or they use a fixed return address. In most cases we can find suitable candidates using regular expressions though, this will be demonstrated later in this section.

@@

@@

In order to find the gadgets we want we can use the tools directly or we can use regular expressions. In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.

@@

The following table describes what regex we can use to find the gadgets needed for the attack.

@@

The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}. Though, then the payload only worked when filling the buffer with 24 Bytes.

@@

The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the ret in main.

@@

The following program is the target of our attack, it uses a command line argument to provide the payload and \Verb+strcpy+ for the buffer overflow, overwriting the return address after the 8 Byte buffer. It is also possible to perform vulnerable input functions like \bltInlineVerb{scanf}.

@@

  \item Specify goal with required program state and instructions
  \item Generate desired list of instructions and arguments (abstract payload/rop chain)

@@

  \item Search gadgets for instructions

@@

  \item Generate payload using the gadgets according to the the abstract payload while making sure gadgets dont interfere with our desired program state. This step can be done using Python which we will show in a later section~\cref{howtopack}

@@

\paragraph{Goal and abstract payload}
After specifying the goal and possibly simplifying it we have to write a list of instructions and arguments that achieve the goal, for this it is favorable to directly use the format of the final payload except for using instructions instead of addresses as this will then allow to simply insert the found gadgets into this abstract payload. For the example in this paper we want to open a shell, for that the simplest way is to execute an execve system call. The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}

@@

\paragraph{Extract and search gadgets}

@@

There segments in a binary can be read only or writable. It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}. However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment.

@@

>>>>

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
BIBLIOGRAFIE NICHT VERGESSEN
Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham~\cref{hshacham} and has become a widely known buffer overflow technique since. It has been developed to circumvent the NX-BIT protection that protects the stack from being executed. At the time of writing this paper modern techniques like stack carnaries and ASLR make these attacks hard and very time consuming on modern systems. That is not to say ASLR and stack canaries can not be broken by bruteforcing or side channels. Since there are millions of running systems with old hard-, firm- and software that is possibly vulnerable to these kinds of attacks it is still relevant to this day. The main idea in ROP is based on chaining return addresses to code just before a return and therefore allowing almost arbitrary cpu instructions to be chained.

\section{Gadgets}
\label{sec:gadgets}
\paragraph{Introduction}
On the x86 architecture the \Verb+ret+ instruction is defined to pop the return instruction pointer from the stack into the \Verb+eip+ register and redirect code execution to that memory address.~\cite{retx86} A ROP gadget consists of a few instructions (usually 1-3) that end on a \bltInlineVerb{ret}.

@@

A gadget can be found by searching for \Verb+0xC3+ Bytes in the program. The instructions before then represent the code code that can be executed by injecting the addresses of these instructions. It is possible to search for gadgets with \Verb+objdump+ or \Verb+hexdump+, however, the tools specifically made for finding ROP gadgets are really easy to use and provide lots of customizability and features for finding the required gadgets. To name a few ROP gadget tools there is \Verb+ropper+, \Verb+ROPgadget+ and \Verb+pwntools+. For this paper the software \Verb+ROPgadget+ has been employed since i found it easy to use. \bltInlineVerb{ROPgadget} can be found in most package managers or can be downloaded directly from \url{https://github.com/JonathanSalwan/ROPgadget}. The gadgets can be extracted from the file with the following command~\cref{dumpallgadgets}. We can then use regular expressions or ROPgadget directly to search for the required gadgets.

@@

These are only 10 Lines out of the 8244 lines found by the tool though and i purposefully filtered out some good and bad ones for demonstration. It is clearly visible that many candidates for ROP can be found, even in a file with a relatively small size of 72 kB. Though most of these gadgets are not all that useful because they often modify a lot of registers, possibly messing up the desired state. In most cases we can find suitable candidates using regular expressions, this will be demonstrated later in this section~\cref{subsec:filtering}.

@@

\label{subsec:filtering}

@@

In order to find the required gadgets we can use the tools directly or we can use regular expressions. In order to make this paper more general and easy to replicate i will be using regular expressions to find the desired gadgets.

@@

The following table describes what regex we can use to find the gadgets required for the attack.

@@

The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s, because of that the buffer has to be filled with 16 Bytes instead of 8 Bytes, this can be turned off with the option \bltInlineVerb{-mpreferred-stack-boundary=2}. Surprisingly the payload then only worked when filling the buffer with 24 Bytes.

@@

The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}.

@@

The following program is the target of our attack, it uses a command line argument to provide the payload and \Verb+strcpy+ for the buffer overflow, overwriting the return address after the 8 Byte buffer.

@@

  \item Specify attack, analyze necessary setup to be done.~\cref{par:goal}

@@

@@

  \item Generate payload with the extracted gadgets based on the specification in step 1.

@@

\paragraph{Specification and abstract payload}
\label{par:goal}
After specifying the goal and possibly simplifying it we have to determine the required program state. For the example in this paper we want to open a shell, for that the simplest way is to execute an \bltInlineVerb{execve} system call. The following program state~\cref{fig:stateforint} has to be achieved so the interrupt \bltInlineVerb{int 0x80} causes a shell to be opened.~\cite{pixis}~\cite{proggen-rop}

@@

\paragraph{Extract gadgets}
The gadgets can be extracted and like described in~\cref{sec:gadgets}

@@

The segments in a binary can be read only or writable. It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}. However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment. The data segment contains static and global variables. Since the target program does not have any global or static variables we can override this segment with arbitrary character sequences. In 

@@

\paragraph{Generating the payload}
With this information we can start to construct the ideal payload, based on the description above and some knowledge about assembly the payload could take the following form. \\
\bltInlineVerb{pop edx | 0x080e5020 | pop eax | "/bin" | pop edx | 0x080e5024 | "//sh" | xor eax, eax | pop edx | 0x080e5028 | mov dword ptr [edx], eax | pop ebx | 0x080e5020 | xor ecx, ecx | xor edx, edx | xor eax, eax | (inc eax) * 11 | int 0x80} \\
When constructing this ideal payload it is important to know that some \bltInlineVerb{string.h} functions use the \bltInlineVerb{0x00} Byte to identify the end of a string. This means that depending on the implementation of the target it is important to not insert any \bltInlineVerb{0x00} Bytes into the payload otherwise the buffer does overflow fully. In most cases we can still write \bltInlineVerb{0x00} Bytes into registers or into memory. This can be accomplished by \bltInlineVerb{xor}'ing a register with itself and then copying that value into a register or into memory.
